---
title: Introduction to `Require`
output: html_document
editor_options: 
  chunk_output_type: console
---

`Require` is a single package that combines features of `base::install.packages`, `base::library`, `base::require`, as well as `pak::pkg_install`, `remotes::install_github`, and `versions::install_version`, plus the snapshotting capabilities of `renv`. It takes its name from the idea that a user could simply have one line named from the `require` function that would load a package, but in this case it will also install the package if necessary. Set it and forget it. This means that even if a user has a dependency that is removed from CRAN ("archived"), the line will still work. Because it can be done in one line, it becomes relatively easy to share, which facilitates, for example, making reprexes for debugging. 


# Principles used in `Require`

`Require` is desinged with features that facilitate running R code that is part of a continuous reproducible workflow, from data-to-decisions. For this to work, all functions called by a user should have a property whereby the initial time they are called does the heavy work, and the subsequent times are sufficiently fast that the user is not forced to skip over lines of code when re-running code. The package, `reproducible`, has a function `Cache` which can convert many function calls to have this property. It does not work well for functions whose objectives are side-effects, like installing and loading packages. `Require` fills this gap.

## Key features

Features include:

1. Fast, parallel installs and downloads;
2. Installs CRAN *even if they have been archived.*;
3. Installs GitHub packages;
4. User can specify which version to install.
5. Local package caching and cloning for fast (re-)installs.
6. Manages (some types of) conflicting package requests, i.e., different GitHub branches.
7. `options`-level control of which packages should be installed from source (see `RequireOptions()`) even if they are being downloaded from a binary repository.

## How it works

`Require` uses `install.packages` internally to install packages. However, it does not let `install.packages` download the packages. Rather, it identifies dependencies recursively, finds out where they are (CRAN, GitHub, Archives, Local), downloads them (or gets from local cache or clones from an specified package library). If `libcurl` is available (assessed via `capabilities("libcurl")`), it will download them in parallel from CRAN-alike repositories. If `sys` is installed, it will download GitHub packages in parallel also. If a user has not set `options("Ncpus")` manually, then it will set that to a value up to 8 for parallel installs of binary and source packages.

## Why these features help teams 

It is common during code development to work in teams, and to be updating package code. This is beneficial whether the team is very tight, all working on exactly the same project, or looser where they only share certain components across diverse projects.

### All working on same project

If the whole team is working on the same "whole" project, then it may be useful to use a "package snapshot" approach, as is used with the `renv` package. `Require` offers similar functionality with the function `pkgSnapshot()`. Using this approach provides a mechanism for each team member to update code, then snapshot the project, commit the snapshot and push to the cloud for the team to share. 

### Diverse projects

However, if a team is more diversified and they are actually sharing the new code, but not the whole project, then project snapshots will be very inefficient and package management must be on a package-by-package case, not the whole project. In other words, the code developer can work on their package, and the various team members will have 2 options of what they might want to do: keep at the bleeding edge or update only if necessary for dependencies. More likely, they will want to have a mixture of these strategies, i.e., bleeding edge with some code, but only if necessary with others. Thus, `Require` offers programmatic control for this. For example 

```{r,eval=FALSE}
library(Require)
Require::Install(
  c("PredictiveEcology/reproducible@development (HEAD)", 
    "PredictiveEcology/SpaDES.core@development (>=2.0.5.9004)")) 
```
will keep the project at the bleeding edge of the development branch of `reproducible`, but will only update if necessary (based on the version needed, expressed by the inequality) for the development branch of `SpaDES.core`. The user doesn't have to make decisions at run time as to whether an update should be made, and for which packages.

# How `Require` differs from other approaches

### Default behaviours different

For packages that aren't yet installed:

| Description                      |  Outcome                                   |
| -------------------------------- | ------------------------------------------ |
| `Install("data.table")`          | `data.table` installed *if not installed*  |
| `install.packages("data.table")` | `data.table` installed                     |
| `pak::pkg_install("data.table")` | `data.table` installed *if not installed*  |
| `renv::install("data.table")`    | `data.table` installed                     |

For packages that are already installed, but old:

| Description                      |  Outcome                                   |
| -------------------------------- | ----------------------------------------------------------------- |
| `Install("data.table")`          | `data.table` not installed                                        |
| `install.packages("data.table")` | `data.table` installed, always, regardless of version             |
| `pak::pkg_install("data.table")` | `data.table` installed, asks user if wants to update if available |
| `renv::install("data.table")`    | `data.table` installed, asks user if wants to update if available |


### Differences and similarities between `pak` and `Require`

\* Indicates that there is an example below.

| Description                      |  `Require`    |   `pak`                               |
| -------------------------------- | :------------------------------: | :-----------------------------------: |
| Archived package*                 | Automatic                                        | Not an option  | 
| Dependency conflicts*   | Yes   |  No (see example below)  |  
| Multiple requests of same package | Resolves by version number specification, or most recent version | Error |
| Control individual package updates | With `HEAD`   |   No   |
| Parallel downloads   |   Yes    |   Yes    |
| Parallel installs    |   Yes    |   Yes    |
| Very clean messaging   |   somewhat, with `options(Require.installPackagesSys = 1)`  | Yes    |  
| Package dependencies     |  `data.table`   |   None |
| Uses local cache   |   Yes   |   Yes   |
| Package updates (default)    | No, unless needed by version number | Yes, prompt user |     
| Package install by version    | Yes | No (only if declared in a `DESCRIPTION` file)  |   
| Package conflict (CRAN & GitHub) | Prefers CRAN, if version requirements met  |  Error  |

### Archived packages

Between mid March 2024 and April 5, 2024, `fastdigest` was taken off CRAN. If this is part of *your* direct dependencies, you can remove it and find an alternative. However, if it is an indirect dependency, you don't have that choice. `Require` will just get the most recent archived copy and the work can continue.

```{r,eval=FALSE,message=FALSE}
Require::Install("fastdigest")

try(pak::pkg_install(c("fastdigest")))
```

### Dependency conflict

When doing code development, it is common to use many `GitHub` packages. Each of these (or their dependencies) may point to one or more `Remotes`, i.e., *other* GitHub packages. 
```{r, eval=FALSE,message=TRUE}

# Fails because of a) packages taken off CRAN & multiple GitHub branches requested
try(pk <- pak::pkg_deps("PredictiveEcology/LandR@LandWeb"))

# Fine
rq <- Require::pkgDep("PredictiveEcology/LandR@LandWeb")

```

The following doesn't work with `pak` because BioSIM is not found. This may be because the package name is not the repository name, but it is not clear from the error message why: 
```{r,eval=TRUE,message=FALSE}
try(gg <- pak::pkg_deps("PredictiveEcology/LandR@development", dependencies = TRUE))
ff <- Require::pkgDep("PredictiveEcology/LandR@development", dependencies = TRUE)
```



### Version requirements determine package installation

1. **Version number requirements** drive package updates. If a user doesn't need an update because version numbers are sufficient, no update will occur.

2. If no version number specification, then installs only occur if package is not present.

3. Multiple simultaneous requests to install a package from what appear to be incompatible sources, will not create a conflict unless version requirements cause the conflict. If version number requirements are not specified, CRAN versions will take precedence.

```{r,eval=TRUE}
# The following has no version specifications, 
#   so CRAN version will be installed or none installed if already installed
Require::Install(c("PredictiveEcology/reproducible@development", "reproducible"))

# The following specifies "HEAD" after the Github package name. This means the 
#   tip of the development branch of reproducible will be installed if not already installed
Require::Install(c("PredictiveEcology/reproducible@development (HEAD)", "reproducible"))

# The following specifies "HEAD" after the package name. This means the 
#   tip of the development branch of reproducible
Require::Install(c("PredictiveEcology/reproducible@development", "reproducible (HEAD)"))

# Not a problem because version number specifies
Require::Install(c("PredictiveEcology/reproducible@modsForLargeArchives (>=2.0.10.9010)",
                   "PredictiveEcology/reproducible (>= 2.0.10)"))

# Even if branch does not exist, if later version requirement specifies a different branch, no error
Require::Install(c("PredictiveEcology/reproducible@modsForLargeArchives (>=2.0.10.9010)",
                   "PredictiveEcology/reproducible@validityTest (>= 2.0.9)"))
```

```{r,eval=TRUE}
try(pak::pkg_install(
  c("PredictiveEcology/reproducible@development", 
    "PredictiveEcology/reproducible")))


## FAILS - can't specify version requirements
try(pak::pkg_install(
    c("PredictiveEcology/reproducible@modsForLargeArchives (>=2.0.10.9010)",
      "PredictiveEcology/reproducible (>= 2.0.10)")))
```

Fails when there are 2 branches within a DESCRIPTION -- wants to downgrade, regardless of order
```{r}
# Installs or does nothing if `reproducible` already installed
twoReproducibles <- c("PredictiveEcology/reproducible@development", 
                      "PredictiveEcology/reproducible")
Require::Install(twoReproducibles)

# Fails because you can't do this
try(pak::pkg_install(twoReproducibles))

```

### Error handling

`pak` uses a different approach to error handling. Sometimes this can be effective. Other times it limits the ability to diagnose errors:
```{r,eval=FALSE}

```

## Binary on Linux

On Linux, users have the ability to install binary packages that are pre-built e.g., from the Posit Package Manager. Sometimes the binary is incompatible with a user's system, even though it is the correct operating system. This occurs generally for several packages, and thus they must be installed from source. `Require` has a function `sourcePkgs()`, which can be informed by `options("Require.spatialPkgs")` and `options("Require.otherPkgs")` that can be set by a user on a package-by-package basis. By default, some are automatically installed from `"source"` because in our experience, they tend to fail if installed from the binary.

```{r,eval=FALSE}
# In this example, it is `terra` that must be installed from source on Linux
if (Require:::isLinux()) {
  Require::setLinuxBinaryRepo()
  pkgs <- c("terra", "PSPclean")
  pkgFullName <- "ianmseddy/PSPclean@development"
  try(remove.packages(pkgs))
  pak::cache_delete() # make sure a locally built one is not present in the cache
  try(pak::pkg_install("ianmseddy/PSPclean@development"))
  # ✔ Loading metadata database ... done                                       
  #                                                                            
  # → Will install 2 packages.
  # → Will download 2 packages with unknown size.
  # + PSPclean   0.1.4.9005 [bld][cmp][dl] (GitHub: fed9253)
  # + terra      1.7-71     [dl] + ✔ libgdal-dev, ✔ gdal-bin, ✔ libgeos-dev, ✔ libproj-dev, ✔ libsqlite3-dev
  # ✔ All system requirements are already installed.
  #   
  # ℹ Getting 2 pkgs with unknown sizes
  # ✔ Got PSPclean 0.1.4.9005 (source) (43.29 kB)                         
  # ✔ Got terra 1.7-71 (x86_64-pc-linux-gnu-ubuntu-22.04) (4.24 MB)      
  # ✔ Downloaded 2 packages (4.28 MB) in 2.9s                
  # ✔ Installed terra 1.7-71  (61ms)                                   
  # ℹ Packaging PSPclean 0.1.4.9005                                    
  # ✔ Packaged PSPclean 0.1.4.9005 (420ms)                              
  # ℹ Building PSPclean 0.1.4.9005                                      
  # ✖ Failed to build PSPclean 0.1.4.9005 (3.7s)                        
  # Error:                                                              
  # ! error in pak subprocess
  # Caused by error in `stop_task_build(state, worker)`:
  # ! Failed to build source package PSPclean.
  # Type .Last.error to see the more details.
  
  
  # Works fine because the `sourcePkgs()`            
  
  try(remove.packages(pkgs))
  Require::clearRequirePackageCache(pkgs, ask = FALSE)
  Require::Install(pkgFullName)
}
```

### Package dependencies

The package dependencies algorithm is slightly different with the default settings. For instance LinkingTo is not used by default in `pak::pkg_dep`
```{r,eval=FALSE}
depPak <- pak::pkg_deps("PredictiveEcology/LandR@LandWeb")
depRequire <- Require::pkgDep("PredictiveEcology/LandR@LandWeb") # Slightly different default in Require

# Same
pakDepsClean <- setdiff(Require::extractPkgName(depPak$ref), Require:::.basePkgs)
requireDepsClean <- setdiff(Require::extractPkgName(depRequire[[1]]), Require:::.basePkgs)
setdiff(pakDepsClean, requireDepsClean)
setdiff(requireDepsClean, pakDepsClean) # differences are from missing LinkingTo for e.g., fasterize

```


## default arguments -- `pkgDep(..., which = XX)` includes `LinkingTo`

`pkgDep`, by default, includes `LinkingTo` as these are required by `Rcpp` if that is required, and so are strictly necessary.
`pak::pkg_deps` does not include `LinkingTo` by default.

## default behaviour -- CRAN-preference

If there is no version specification, `pkgDep` prefers CRAN packages when there are multiple pointers to a package. Thus, even though a package may have a `Remotes` field pointint to e.g., `PredictiveEcology/SpaDES.tools@development`, if there is a recursive dependency within that package that specifies `SpaDES.tools` without a `Remotes` field, then `pkgDep` will return the `CRAN` version. If a user wants to override this behaviour, then the user can specify a version requirement that can only be satisfited with the `Remotes` option. Then `pkgDep` will take that.

`pak::pkg_deps` prefers the top-level specification, i.e., the non-recursive `Remotes` field will be returned, even if the same package is also specified within a recursive dependency without a `Remotes` field, i.e, if a recursive dependency points the CRAN package, it will not return that version of the dependency.


### `pak` fails for packages on GitHub that are not same name as Git Repo in Remotes

```{r,eval=FALSE}
gg <- pak::pkg_deps("PredictiveEcology/LandR@development", dependencies = TRUE)
# Error:                                                                                  
# ! error in pak subprocess
# Caused by error: 
# ! Could not solve package dependencies:
# * PredictiveEcology/LandR@development: Can't install dependency BioSIM
# * BioSIM: Can't find package called BioSIM.
# Type .Last.error to see the more details.
ff <- Require::pkgDep("PredictiveEcology/LandR@development", dependencies = TRUE)
# $`PredictiveEcology/LandR@development`
#  [1] "BIEN"                                                  "RNCan/BioSimClient_R"
#  [3] "DBI (>= 0.8)"                                          "ENMeval"
#  [5] "Formula"                                               "Hmisc"
#  [7] "PredictiveEcology/LandR@development"                   "MatrixModels"
#  ...
```

# `renv` and `Require`

## Managing projects during development

`renv` has a concept of a lockfile. This lockfile records a specific version of a package. If the current installed version of a package is different from the lockfile (e.g., I am the developer and I increment the local version), `renv` will attempt to revert the local changes (with prompt to confirm) *unless* the local package is installed from a cloud repository (e.g., GitHub), and a `snapshot` is taken. This sequence is largely incompatible with `pkgload::load_all()` or `devtools::install()`, as these do not record "where" to get the current version from. Thus, the `renv` sequence can be quite time consuming (1-2 minutes, instead of 1 second with `pkgload::load_all()`).

`Require` does not attempt to update anything unless required by a package. Thus, this issue never comes up. If and when it is important to "snapshot", then `pkgSnapshot` or `pkgSnapshot2` can be used. 

## Using `DESCRIPTION` file to maintain minimum versions

During a project, a user can build and maintain and "project-level" DESCRIPTION file, which can be useful for a `renv` managed project. This approach does not, however, automatically detect minimum version changes or GitHub branch changes (`renv::status` does not recognize these). In order for a user to inherit the correct requirements, a manual [`renv::install` must be used](https://github.com/rstudio/renv/issues/233#issuecomment-1530134112). For even moderate sized projects, this can take over 20 seconds. 

`Require` does not need a lockfile; package violations are found on the fly. 
